---
title: "p8105_hw2_jz3571"
author: "Junyan Zhu"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

```{r}
nyc_df = read_csv('./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv') %>% 
  janitor::clean_names() %>% 
  select(line:station_longitude, starts_with("route"), entry, exit_only, vending, entrance_type, ada) %>% 
  mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
  
```

The date has `r nrow(nyc_df)` rows and `r ncol(nyc_df)` columns and it contains variables: `r names(nyc_df)`. The above date cleaning steps include selecting required variables, converting route 8-11 into character variables and converting entry variable from chatacter variable to logic variable. These data are not tidy yet.

- Number of Distint Stations

```{r}
nyc_df %>% 
  select(station_name, line) %>% 
  distinct %>% 
  nrow()
```

There are 465  distinct stations.

- Number of Ada Compliant Stations

```{r}
nyc_df %>% 
  filter(ada == 'TRUE') %>% 
  select(station_name, line) %>% 
  distinct %>% 
  nrow()
```

There are 84 stations are ADA compliant


- Number of stations without vending alowing entrance

```{r}
nyc_df %>% 
  filter(vending == 'NO') %>% 
  pull(entry) %>% 
  mean()
```

There are 37.7% of the entrances/exits without vending allow entrance.

- Make Route Number Distinct Variable

```{r}
nyc_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route"
  ) %>% 
  filter(route == "A") %>% 
  distinct(station_name, line) %>% 
  nrow()

nyc_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route"
  ) %>% 
  filter(route == "A", ada == "TRUE") %>% 
  distinct(station_name, line) %>% 
  nrow()
```

There are 60 distinct stations serve the A train, of which 17 are ada compliant

# Problem 2

1) Clean MR.Trash Wheel Data

```{r}
mr_df = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") %>% 
  janitor::clean_names()
```

```{r}
tidy_mr_df = mr_df %>% 
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(vessel_type = "Mr.Trash Wheel") %>%
  mutate(dumpster = as.integer(dumpster))%>%
  drop_na(dumpster)
```

2) Clean Professor Trash Wheel Data

```{r}
pr_df = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") %>% 
  janitor::clean_names()
```

```{r}
tidy_pr_df = pr_df %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(vessel_type = "Professor Trash Wheel") %>% 
  mutate(year = as.character(year))
```

3) Bind Two Together

```{r}
bind_df = bind_rows(tidy_mr_df, tidy_pr_df)
```

After combining two datasets, we have `r nrow(bind_df)` rows and `r ncol(bind_df)`
columns. `vessel_type` is an added variable to distinguish the two different wheels used. Other key variables include dumspter number, month, year, date, weighted trash in tons, amount of different trash, and homes powered by trash-generated electricity.

- Total weigh of trash collected to Professor Trash Wheel

```{r}
bind_df %>% 
  filter(vessel_type == "Professor Trash Wheel") %>% 
  select(weight_tons) %>% 
  sum()
```

- Total number of sports balls collected by Mr. Trash Wheel

```{r}
bind_df %>% 
  filter(vessel_type == "Mr.Trash Wheel") %>% 
  select(sports_balls) %>% 
  sum()
```

# Problem 3

1) Clean plol_month data

```{r}
pols_month = read_csv('./data/pols_month.csv') %>% 
  janitor::clean_names()
```

```{r}
tidy_month = pols_month %>% 
  separate(mon, into = c('year', 'month', 'date'), sep = '-') %>% 
  mutate(month = month.name[as.numeric(month)], year= as.integer(year)) %>% 
  mutate(president = ifelse(prez_gop >= 1, "Republican", 'Democract')) %>% 
  select(-prez_gop, -prez_dem, -date)
```

2) Clean snp data

```{r}
snp_df = read_csv('./data/snp.csv') %>% 
  janitor::clean_names()
```

```{r}
tidy_snp = snp_df %>% 
  separate(date, into = c('month','date','year'), sep = '/') %>% 
  mutate(month = as.integer(month), date = as.integer(date), year = as.integer(year)) %>% 
  mutate(year = ifelse(year <= 15, year+2000, year+1900)) %>% 
  arrange(year, month) %>% 
  select(year, month, close)
```

3) Clean unemployment data

```{r}
unemp_df = read_csv('./data/unemployment.csv') %>% 
  janitor::clean_names()
```

```{r}
tidy_unemp = unemp_df %>% 
  pivot_longer(2:13,
               names_to = "month",
               values_to = "unemployment_rate") %>% 
  mutate(month = recode(month,
                        'jan' = 'January',
                        'feb' = 'February', 
                        'mar' = 'March', 
                        'apr' = 'April', 
                        'may' = 'May', 
                        'jun' = 'June', 
                        'jul' = 'July', 
                        'aug' = 'August', 
                        'sep' = "September", 
                        'oct' = 'October', 
                        'nov' = 'November', 
                        'dec' = 'December'), 
         year = as.integer(year))
```



